Bacterial Microbiome and Obesity Meta-analysis Supplemental

### 2016-03-23 Sze_ObesityMicrobiomeMetaNoMetaHit_PeerJ_2016.Rmd
manuscript and Suppl update

demographics <- read.csv("results/tables/denovodemographicsTable.csv")
OverallPValue <- read.csv("results/tables/denovoOverallPTable.csv")
ShannonRRTable <- read.csv("results/tables/denovoShannonRRTable.csv")
BFRatioRRTable <- read.csv("results/tables/denovoBFRatioRRTable.csv")
ZScoreCombinedData <- read.csv("results/tables/denovoCombinedData.csv")
ZScoreTurnbaughData <- read.csv("results/tables/denovoTurnbaughData.csv")
AUCRFData <- read.csv("results/tables/denovoAUCRFDataTable.csv")
PowerTable <- read.csv("results/tables/denovoPowerTable.csv")

@@

```{r install_needed_packages, echo=FALSE, message=FALSE, warning=FALSE}
deps = c("knitr","vegan","epiR","AUCRF","meta","metafor","ggplot2", "plyr", "reshape2", "RColorBrewer", "scales", "gridExtra");
for (dep in deps){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}

@@

**Table 3. Summary of P-values for Measurements of Interest for each Individual Study for Obese versus Normal**

```{r Table_3, echo=FALSE, message=FALSE, warning=FALSE}

Study <- as.character(OverallPValue$X)
Bacteroidetes <- format(round(OverallPValue$Bacteroidetes, 3), nsmall=3)
Firmicutes <- format(round(OverallPValue$Firmicutes, 3), nsmall=3)
BFRatio <- format(round(OverallPValue$BFRatio, 3), nsmall=3)
Shannon <- format(round(OverallPValue$Shannon, 3), nsmall=3)
OTURich <- format(round(OverallPValue$OTURich, 3), nsmall=3)
Evenness <- format(round(OverallPValue$Evenness, 3), nsmall=3)
BrayC <- format(round(OverallPValue$BrayC, 3), nsmall=3)

table3 <- cbind(Study, Bacteroidetes, Firmicutes, BFRatio, Shannon, OTURich, Evenness, BrayC)
colnames(table3) <- c("Study", "Bacteroidetes", "Firmicutes", "B/F Ratio", "Shannon Diversity", "OTU Richness", "Evenness", "Bray Curtis")

@@

**Table 4. Summary of Power Calculations for each Individual for Shannon Diversity and B/F ratio for Non-obese versus Obese**

@@

Study <- as.character(PowerTable$Study)
PowerH <- PowerTable$StudyPowerH
PowerBF <- PowerTable$StudyPowerBF

table4 <- cbind(Study, PowerH, PowerBF)
colnames(table4) <- c("Study", "Shannon Diversity Power (%)", "B/F Ratio Power (%)")

kable(table4, format = "markdown", align = 'c', padding = 2)
```

@@

#To create a graph visualize the data
forest(res, at=log(c(0.3, 1, 5)), atransf=exp, xlab="RR of Obesity in high versus low Shannon diversity based on Median", cex=0.8)
par(font=2) #Switch to bold

@@

#To create a graph visualize the data
forest(res2, at=log(c(0.1, 1, 3)), atransf=exp, xlab="RR of Obesity in high versus low B/F ratio based on Median", cex=0.8)
par(font=2) #Switch to bold

@@

```{r Figure_4, echo=FALSE, fig.path="results/figures/"}
par(mfrow=c(1,2))
funnel(res, main="A")
funnel(res2, main="B")
```

**Figure 4: Funnel Plot of the Shannon Diversity and Bacteroidetes/Firmicutes Ratio Relative Risk.**  **A)** Overall there does not seem to be any bias associated with the studies selected for the Shannon Diversity analysis with studies falling on either side of the predicted value and those with smaller total n falling further away from this. **B)** For the Bacteroidetes/Firmicutes ratio there does not seem to be any associated bias. The overall pattern is similar to that observed for the Shannon Diversity analysis.

```{r Figure_5, echo=FALSE, message=FALSE, warning=FALSE, fig.path="results/figures/"}
# Section A
p1 <- ggplot(group.data.test, aes(obese, ZH))
p2 <- p1 + geom_boxplot(aes(fill = obese)) + scale_fill_brewer(palette="Set1")
p3 <- p2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
p4 <- p3 + xlab("Obese Category") + ylab("Z-score Normalized H")
p5 <- p4 + guides(fill=guide_legend(title="Obese")) + theme(legend.title = element_text(size=8, face="bold"))
p6 <- p5 + theme(axis.line=element_line(colour="black"),
                 panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.border = element_blank(),
                 panel.background = element_blank())
p7 <- p6 + ggtitle('A') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

# Section B
q1 <- ggplot(group.data.test, aes(BMI.class, ZH))
q2 <- q1 + geom_boxplot(aes(fill = BMI.class)) + scale_fill_brewer(palette="Set1")
q3 <- q2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
q4 <- q3 + xlab("BMI Classification") + ylab("Z-score Normalized H")
q5 <- q4 + guides(fill=guide_legend(title="BMI Class")) + theme(legend.title = element_text(size=8, face="bold"))
q6 <- q5 + theme(axis.line=element_line(colour="black"),
                 panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.border = element_blank(),
                 panel.background = element_blank())
q7 <- q6 + ggtitle('B') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

# Section C
bf1 <- ggplot(group.data.test, aes(obese, ZLogBF))
bf2 <- bf1 + geom_boxplot(aes(fill = obese)) + scale_fill_brewer(palette="Set1")
bf3 <- bf2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
bf4 <- bf3 + xlab("Obese Category") + ylab("Z-score Normalized Log B/F Ratio")
bf5 <- bf4 + guides(fill=guide_legend(title="Obese")) + theme(legend.title = element_text(size=8, face="bold"))
bf6 <- bf5 + theme(axis.line=element_line(colour="black"),
                   panel.grid.major = element_blank(),
                   panel.grid.minor = element_blank(),
                   panel.border = element_blank(),
                   panel.background = element_blank())
bf7 <- bf6 + ggtitle('C') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

# Section D
qbf1 <- ggplot(group.data.test, aes(BMI.class, ZLogBF))
qbf2 <- qbf1 + geom_boxplot(aes(fill = BMI.class)) + scale_fill_brewer(palette="Set1")
qbf3 <- qbf2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
qbf4 <- qbf3 + xlab("BMI Classification") + ylab("Z-score Normalized Log B/F Ratio")
qbf5 <- qbf4 + guides(fill=guide_legend(title="BMI Class")) + theme(legend.title = element_text(size=8, face="bold"))
qbf6 <- qbf5 + theme(axis.line=element_line(colour="black"),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),
                     panel.border = element_blank(),
                     panel.background = element_blank())
qbf7 <- qbf6 + ggtitle('D') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

@@

**Figure 5: Pooled Analysis of Shannon Diversity and B/F ratio by either Obesity Status or BMI Groups.**  **A)** Z-score Normalized Shannon Diversity and Obesity. There was a significant difference between non-obese (n = `r length(which(group.data.test$obese == "No"))`) and obese groups (n = `r length(which(group.data.test$obese == "Yes"))`) (P-value = `r signif(combZHTtest$p.value, 3)`). **B)** Z-score Normalized Shannon Diversity by BMI Group. Total n for lean, overweight, and obese groups are `r length(which(group.data.test$BMI.class == "Lean"))`, `r length(which(group.data.test$BMI.class == "Overweight"))`, `r length(which(group.data.test$BMI.class == "Obese"))` respectively. Using an ANOVA with tukey post-hoc testing there was a significant diference between the overweight and obese group (P-value = `r signif(combZHTukeyFinal[3, 4], 3)`) but no difference between lean and obese group (P-value = `r signif(combZHTukeyFinal[2, 4], 3)`). **C)** Z-score Normalized B/F Ratio and Obesity. There was no significant difference between the non-obese (n = `r length(which(group.data.test$obese == "No"))`) and obese (n = `r length(which(group.data.test$obese == "Yes"))`) group based on B/F ratio (P-value = `r signif(combZLogBFTtest$p.value, 3)`). **D)** Z-score Normalized B/F Ratio by BMI Group. Total n for lean, overweight, and obese groups are `r length(which(group.data.test$BMI.class == "Lean"))`, `r length(which(group.data.test$BMI.class == "Overweight"))`, `r length(which(group.data.test$BMI.class == "Obese"))` respectively. There was no significant difference between any of the three groups based on ANOVA with Tukey post-hoc testing (P-value = `r signif(combZLogBFANOVA[1, 5], 3)`).

>>>>

demographics <- read.csv("results/tables/denovodemographicsTable.csv")
OverallPValue <- read.csv("results/tables/denovoOverallPTable.csv")
ShannonRRTable <- read.csv("results/tables/denovoShannonRRTable.csv")
BFRatioRRTable <- read.csv("results/tables/denovoBFRatioRRTable.csv")
ZScoreCombinedData <- read.csv("results/tables/denovoCombinedData.csv")
ZScoreTurnbaughData <- read.csv("results/tables/denovoTurnbaughData.csv")
AUCRFData <- read.csv("results/tables/denovoAUCRFDataTable.csv")
PowerTable <- read.csv("results/tables/denovoPowerTable.csv")
ModelledPower1PerDiff <- read.csv("results/tables/denovoModelled1PerDiff.csv")
ModelledPower5PerDiff <- read.csv("results/tables/denovoModelled5PerDiff.csv")
ModelledPower10PerDiff <- read.csv("results/tables/denovoModelled10PerDiff.csv")
ModelledPower15PerDiff <- read.csv("results/tables/denovoModelled15PerDiff.csv")
ActualDiff <- read.csv("results/tables/denovoActualDiff.csv")
PowerModelHRR <- read.csv("results/tables/denovoPowerModelHRR.csv")
PowerModelBFRR <- read.csv("results/tables/denovoPowerModelBFRR.csv")
SimulatedN_H <- read.csv("results/tables/denovoSimulatedN_H.csv")
SimulatedN_BF <- read.csv("results/tables/denovoSimulatedN_BF.csv")
RRSimNNeeded <- read.csv("results/tables/denovoRRSimNNeeded.csv")

@@

```{r install_needed_packages, echo=FALSE, message=FALSE, warning=FALSE}
deps = c("knitr","vegan","epiR","AUCRF","metafor","ggplot2", "plyr", "reshape2", "RColorBrewer", "scales", "gridExtra", "pwr", "statmod");
for (dep in deps){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}

@@

**Table 3. Summary of Power Calculation Simulations for Shannon Diversity for Non-obese versus Obese**

```{r Table_3, echo=FALSE, message=FALSE, warning=FALSE}
ObeseTotal <- ShannonRRTable$tposH + ShannonRRTable$cposH
ContTotal <- ShannonRRTable$tnegH + ShannonRRTable$cnegH

Study <- paste(as.character(PowerTable$Study), " (", ContTotal, "|", ObeseTotal, ")", sep="")
ActualDiffH <- format(round(ActualDiff$H, 3), nsmall=3)
Shannon <- format(round(OverallPValue$Shannon, 2), nsmall=2)
OneP <- format(round(ModelledPower1PerDiff$H, 3), nsmall=3)
FiveP <- format(round(ModelledPower5PerDiff$H, 3), nsmall=3)
TenP <- format(round(ModelledPower10PerDiff$H, 3), nsmall=3)
FifteenP <- format(round(ModelledPower15PerDiff$H, 3), nsmall=3)

table3 <- cbind(Study, ActualDiffH, Shannon, OneP, FiveP, TenP, FifteenP)
colnames(table3) <- c("Study (Non-Obese|Obese)", "Actual Percent Difference", "P-value", "Estimated Power for 1% Difference", "Estimated Power for 5% Difference", "Estimated Power for 10% Difference", "Estimated Power for 15% Difference")

@@

**Table 4. Summary of Power Calculation Simulations for B/F Ratio for Non-obese versus Obese**

@@

ObeseTotal <- ShannonRRTable$tposH + ShannonRRTable$cposH
ContTotal <- ShannonRRTable$tnegH + ShannonRRTable$cnegH

Study <- paste(as.character(PowerTable$Study), " (", ContTotal, "|", ObeseTotal, ")", sep="") 
ActualDiffBF <- format(round(ActualDiff$BFratio, 3), nsmall=3)
BF <- format(round(OverallPValue$BFRatio, 2), nsmall=2)
OneP <- format(round(ModelledPower1PerDiff$BFratio, 3), nsmall=3)
FiveP <- format(round(ModelledPower5PerDiff$BFratio, 3), nsmall=3)
TenP <- format(round(ModelledPower10PerDiff$BFratio, 3), nsmall=3)
FifteenP <- format(round(ModelledPower15PerDiff$BFratio, 3), nsmall=3)

table4 <- cbind(Study, ActualDiffBF, BF, OneP, FiveP, TenP, FifteenP)
colnames(table4) <- c("Study (Non-Obese|Obese)", "Actual Percent Difference", "P-value", "Estimated Power for 1% Difference", "Estimated Power for 5% Difference", "Estimated Power for 10% Difference", "Estimated Power for 15% Difference")

kable(table4, format = "markdown", align = 'c', padding = 2)

**Table 5. Summary of Sample Size Needed Simulations for Shannon Diversity for Non-obese versus Obese**

```{r Table_5, echo=FALSE, message=FALSE, warning=FALSE}

ObeseTotal <- ShannonRRTable$tposH + ShannonRRTable$cposH
ContTotal <- ShannonRRTable$tnegH + ShannonRRTable$cnegH

Study <- paste(as.character(PowerTable$Study), " (", ContTotal, "|", ObeseTotal, ")", sep="")

NNeeded <- format(round(SimulatedN_H[, -1],0), nsmall=0)
table5 <- cbind(Study, NNeeded)
colnames(table5) <- c("Study (Non-Obese|Obese)", "Actual n/Group Needed", "Estimated n/Group for 1% Difference", "Estimated n/Group for 5% Difference", "Estimated n/Group for 10% Difference", "Estimated n/Group for 15% Difference")

kable(table5, format = "markdown", align = 'c', padding = 2)
```

**Table 6. Summary of Sample Size Needed Simulations for B/F Ratio for Non-obese versus Obese**

```{r Table_6, echo=FALSE, message=FALSE, warning=FALSE}

ObeseTotal <- ShannonRRTable$tposH + ShannonRRTable$cposH
ContTotal <- ShannonRRTable$tnegH + ShannonRRTable$cnegH

Study <- paste(as.character(PowerTable$Study), " (", ContTotal, "|", ObeseTotal, ")", sep="")

NNeeded <- format(round(SimulatedN_BF[, -1],0), nsmall=0)
table6 <- cbind(Study, NNeeded)
colnames(table6) <- c("Study (Non-Obese|Obese)", "Actual n/Group Needed", "Estimated n/Group for 1% Difference", "Estimated n/Group for 5% Difference", "Estimated n/Group for 10% Difference", "Estimated n/Group for 15% Difference")

kable(table6, format = "markdown", align = 'c', padding = 2)
```

@@

#To create a graph visualize the data
forest(res, at=log(c(0.3, 1, 5)), atransf=exp, mlab= "Pooled Relative Risk", xlab="RR of Obesity in high versus low Shannon diversity based on Median", cex=0.8)
par(font=2) #Switch to bold

@@

#To create a graph visualize the data
forest(res2, at=log(c(0.1, 1, 3)), atransf=exp, mlab= "Pooled Relative Risk", xlab="RR of Obesity in high versus low B/F ratio based on Median", cex=0.8)
par(font=2) #Switch to bold

@@

```{r Figure_4, echo=FALSE, message=FALSE, warning=FALSE, fig.path="results/figures/"}
# Section A
p1 <- ggplot(group.data.test, aes(obese, ZH))
p2 <- p1 + geom_boxplot(aes(fill = obese)) + scale_fill_brewer(palette="Set1")
p3 <- p2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
p4 <- p3 + xlab("") + ylab("Z-score Normalized H") + scale_x_discrete(breaks=c("No", "Yes"), labels=c("Non-Obese", "Obese"))
p5 <- p4 + guides(fill=guide_legend(title="Obese")) + theme(legend.title = element_text(size=8, face="bold"))
p6 <- p5 + theme(axis.line=element_line(colour="black"),
                 panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.border = element_blank(),
                 panel.background = element_blank())
p7 <- p6 + ggtitle('A') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

# Section B
q1 <- ggplot(group.data.test, aes(BMI.class, ZH))
q2 <- q1 + geom_boxplot(aes(fill = BMI.class)) + scale_fill_brewer(palette="Set1")
q3 <- q2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
q4 <- q3 + xlab("") + ylab("Z-score Normalized H")
q5 <- q4 + guides(fill=guide_legend(title="BMI Class")) + theme(legend.title = element_text(size=8, face="bold"))
q6 <- q5 + theme(axis.line=element_line(colour="black"),
                 panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.border = element_blank(),
                 panel.background = element_blank())
q7 <- q6 + ggtitle('B') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

# Section C
bf1 <- ggplot(group.data.test, aes(obese, ZLogBF))
bf2 <- bf1 + geom_boxplot(aes(fill = obese)) + scale_fill_brewer(palette="Set1")
bf3 <- bf2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
bf4 <- bf3 + xlab("") + ylab("Z-score Normalized Log B/F Ratio") + 
  scale_x_discrete(breaks=c("No", "Yes"), labels=c("Non-Obese", "Obese"))
bf5 <- bf4 + guides(fill=guide_legend(title="Obese")) + theme(legend.title = element_text(size=8, face="bold"))
bf6 <- bf5 + theme(axis.line=element_line(colour="black"),
                   panel.grid.major = element_blank(),
                   panel.grid.minor = element_blank(),
                   panel.border = element_blank(),
                   panel.background = element_blank())
bf7 <- bf6 + ggtitle('C') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

# Section D
qbf1 <- ggplot(group.data.test, aes(BMI.class, ZLogBF))
qbf2 <- qbf1 + geom_boxplot(aes(fill = BMI.class)) + scale_fill_brewer(palette="Set1")
qbf3 <- qbf2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
qbf4 <- qbf3 + xlab("") + ylab("Z-score Normalized Log B/F Ratio")
qbf5 <- qbf4 + guides(fill=guide_legend(title="BMI Class")) + theme(legend.title = element_text(size=8, face="bold"))
qbf6 <- qbf5 + theme(axis.line=element_line(colour="black"),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),
                     panel.border = element_blank(),
                     panel.background = element_blank())
qbf7 <- qbf6 + ggtitle('D') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

@@

**Figure 4: Pooled Analysis of Shannon Diversity and B/F ratio by either Obesity Status or BMI Groups.**  **A)** Z-score Normalized Shannon Diversity and Obesity. There was a significant difference between non-obese (n = `r length(which(group.data.test$obese == "No"))`) and obese groups (n = `r length(which(group.data.test$obese == "Yes"))`) (P-value = `r signif(combZHTtest$p.value, 3)`). **B)** Z-score Normalized Shannon Diversity by BMI Group. Total n for lean, overweight, and obese groups are `r length(which(group.data.test$BMI.class == "Lean"))`, `r length(which(group.data.test$BMI.class == "Overweight"))`, `r length(which(group.data.test$BMI.class == "Obese"))` respectively. Using an ANOVA with tukey post-hoc testing there was a significant diference between the overweight and obese group (P-value = `r signif(combZHTukeyFinal[3, 4], 3)`) but no difference between lean and obese group (P-value = `r signif(combZHTukeyFinal[2, 4], 3)`). **C)** Z-score Normalized B/F Ratio and Obesity. There was no significant difference between the non-obese (n = `r length(which(group.data.test$obese == "No"))`) and obese (n = `r length(which(group.data.test$obese == "Yes"))`) group based on B/F ratio (P-value = `r signif(combZLogBFTtest$p.value, 3)`). **D)** Z-score Normalized B/F Ratio by BMI Group. Total n for lean, overweight, and obese groups are `r length(which(group.data.test$BMI.class == "Lean"))`, `r length(which(group.data.test$BMI.class == "Overweight"))`, `r length(which(group.data.test$BMI.class == "Obese"))` respectively. There was no significant difference between any of the three groups based on ANOVA with Tukey post-hoc testing (P-value = `r signif(combZLogBFANOVA[1, 5], 3)`).

### 2016-03-23 microbiomeAndobesitySupplemental.Rmd
manuscript and Suppl update

@@

@@

**Table 1. Summary of BLAST alignment of Representative OTU for Clostridiales, Lachnospiraceae, and Ruminococcaceae classification OTUs**

```{r Table_1, echo=FALSE, message=FALSE, warning=FALSE}

@@

table1 <- cbind(Study, VariableRegion, ClosMostPrev, ClostHighMatch, LachMostPrev, LachHighMatch, RumiMostPrev, RumiHighMatch)
colnames(table1) <- c("Study", "Hypervariable Region", "Clostridiales OTU Most Prevalent Match", "Clostridiales OTU Highest Match (Query Cover|Identity)", "Lachnospiraceae OTU Most Prevalent Match", "Lachnospiraceae OTU Highest Match (Query Cover|Identity)", "Ruminococcaceae OTU Most Prevalent Match", "Ruminococcaceae OTU Highest Match (Query Cover|Identity)")

kable(table1, format = "markdown", align = 'c', padding = 2)
```

>>>>

```{r Loading_data_tables, echo=FALSE, message=FALSE, warning=FALSE}

OverallPValue <- read.csv("results/tables/denovoOverallPTable.csv")
ShannonRRTable <- read.csv("results/tables/denovoShannonRRTable.csv")
BFRatioRRTable <- read.csv("results/tables/denovoBFRatioRRTable.csv")

```

@@

**Table S1. Summary of P-values for Measurements of Interest for each Individual Study for Obese versus Normal**

```{r Table_S1, echo=FALSE, message=FALSE, warning=FALSE}

Study <- as.character(OverallPValue$X)
Bacteroidetes <- format(round(OverallPValue$Bacteroidetes, 2), nsmall=2)
Firmicutes <- format(round(OverallPValue$Firmicutes, 2), nsmall=2)
BFRatio <- format(round(OverallPValue$BFRatio, 2), nsmall=2)
Shannon <- format(round(OverallPValue$Shannon, 2), nsmall=2)
OTURich <- format(round(OverallPValue$OTURich, 2), nsmall=2)
Evenness <- format(round(OverallPValue$Evenness, 2), nsmall=2)
BrayC <- format(round(OverallPValue$BrayC, 2), nsmall=2)

tableS1 <- cbind(Study, Bacteroidetes, Firmicutes, BFRatio, Shannon, OTURich, Evenness, BrayC)
colnames(tableS1) <- c("Study", "Bacteroidetes", "Firmicutes", "B/F Ratio", "Shannon Diversity", "OTU Richness", "Evenness", "Bray Curtis")

kable(tableS1, format = "markdown", align = 'c', padding = 2)

```

@@

**Table 2. Summary of BLAST alignment of Representative OTU for Clostridiales, Lachnospiraceae, and Ruminococcaceae classification OTUs**

```{r Table_S2, echo=FALSE, message=FALSE, warning=FALSE}

@@

tableS2 <- cbind(Study, VariableRegion, ClosMostPrev, ClostHighMatch, LachMostPrev, LachHighMatch, RumiMostPrev, RumiHighMatch)
colnames(tableS2) <- c("Study", "Hypervariable Region", "Clostridiales OTU Most Prevalent Match", "Clostridiales OTU Highest Match (Query Cover|Identity)", "Lachnospiraceae OTU Most Prevalent Match", "Lachnospiraceae OTU Highest Match (Query Cover|Identity)", "Ruminococcaceae OTU Most Prevalent Match", "Ruminococcaceae OTU Highest Match (Query Cover|Identity)")

kable(tableS2, format = "markdown", align = 'c', padding = 2)
```

**Table S3. Top Sequence Similarity by Variable Region Based on the Representative OTU for Clostridiales, Lachnospiraceae, and Ruminococcaceae**

```{r Table_S3, echo=FALSE, message=FALSE, warning=FALSE}

VariableRegion <- c("V1-V2 vs V2", "V1-V3", "V4")

TopHitClos <- c("None", "50.3|39.1", "89.4|2.8")
AverageClos <- c("None", "48.0|42.4", "87.7|2.4")
TopHitLachno <- c("TBC", "40.1|47.9", "92.9|0.8")
AverageLachno <- c("TBC", "35.9|52.5", "85.8|5.6")
TopHitRumi <- c("TBC", "30.4|68.4", "86.6|0.8")
AverageRumi <- c("TBC", "29.2|69.1", "81.6|6.6")

tableS3 <- cbind(VariableRegion, TopHitClos, AverageClos, TopHitLachno, AverageLachno, TopHitRumi, AverageRumi)
colnames(tableS3) <- c("Variable Region", "Top Similarity Colstridiales (Id|Gaps)", "Average Similarity Colstridiales (Id|Gaps)", "Top Similarity Lachnospiraceae (Id|Gaps)", "Average Similarity Lachnospiraceae (Id|Gaps)", "Top Similarity Ruminococcaceae (Id|Gaps)", "Average Similarity Ruminococcaceae (Id|Gaps)")

kable(tableS3, format = "markdown", align = 'c', padding = 2)

```

```{r Figure_S4, echo=FALSE, message=FALSE, warning=FALSE, fig.path="results/figures/"}

library(metafor)

# Shannon Diversity Data for Funnel Plot
StudyNo <- as.numeric(ShannonRRTable$X)
Study <- as.character(ShannonRRTable$Study)
Year <- c("2015", "2015", "2014", "2014", "2012", "2011", "2012", "2009")
Total <- c("n=172", "n=63", "n=507", "n=30", "n=200", "n=256", "n=63", "n=146")
tpos <- as.numeric(ShannonRRTable$tposH) #obese in low diversity group
tneg <- as.numeric(ShannonRRTable$tnegH) #nonobese in low diversity group
cpos <- as.numeric(ShannonRRTable$cposH) #obese in high diversity group
cneg <- as.numeric(ShannonRRTable$cnegH) #nonobese in high diversity group
RR <- as.numeric(ShannonRRTable$RRH)
low <- as.numeric(ShannonRRTable$lowH)
high <- as.numeric(ShannonRRTable$highH)

numeric.data <-as.data.frame(cbind(tpos, tneg, cpos, cneg))
raw.overall <- as.data.frame(cbind(StudyNo, Study,Year, Total, numeric.data))

dat <- escalc(measure = "RR", ai = tpos, bi = tneg, ci = cpos, di = cneg, data=raw.overall, append=TRUE)
k <- length(dat$StudyNo)
dat.fm <- data.frame(study=factor(rep(1:k, each = 4)))
dat.fm$grp <- factor(rep(c("T", "T", "C", "C"), k), levels = c("T", "C"))
dat.fm$out <- factor(rep(c("+", "-", "+", "-"), k), levels = c("+", "-"))
dat.fm$freq <- with(raw.overall, c(rbind(tpos, tneg, cpos, cneg)))

#Used to generate the RE Model estimate based on all the studies
res <- rma(ai=tpos, bi=tneg, ci=cpos, di=cneg, data=raw.overall, measure="RR", slab=paste(Study, Year, Total, sep=", "), method="REML")

# B/F ratio data used to generate Funnel Plot
StudyNo <- as.numeric(BFRatioRRTable$X)
Study <- as.character(BFRatioRRTable$Study)
Year <- c("2015", "2015", "2014", "2014", "2012", "2011", "2012", "2009")
Total <- c("n=172", "n=63", "n=507", "n=30", "n=200", "n=256", "n=63", "n=146")
tpos <- as.numeric(BFRatioRRTable$tposBF) #obese in low diversity group
tneg <- as.numeric(BFRatioRRTable$tnegBF) #nonobese in low diversity group
cpos <- as.numeric(BFRatioRRTable$cposBF) #obese in high diversity group
cneg <- as.numeric(BFRatioRRTable$cnegBF) #nonobese in high diversity group
RR <- as.numeric(BFRatioRRTable$RRBF)
low <- as.numeric(BFRatioRRTable$lowBF)
high <- as.numeric(BFRatioRRTable$highBF)

numeric.data <-as.data.frame(cbind(tpos, tneg, cpos, cneg))
raw.overall <- as.data.frame(cbind(StudyNo, Study,Year, Total, numeric.data))

dat <- escalc(measure = "RR", ai = tpos, bi = tneg, ci = cpos, di = cneg, data=raw.overall, append=TRUE)
k <- length(dat$StudyNo)
dat.fm <- data.frame(study=factor(rep(1:k, each = 4)))
dat.fm$grp <- factor(rep(c("T", "T", "C", "C"), k), levels = c("T", "C"))
dat.fm$out <- factor(rep(c("+", "-", "+", "-"), k), levels = c("+", "-"))
dat.fm$freq <- with(raw.overall, c(rbind(tpos, tneg, cpos, cneg)))

#Used to generate the RE Model estimate based on all the studies
res2 <- rma(ai=tpos, bi=tneg, ci=cpos, di=cneg, data=raw.overall, measure="RR", slab=paste(Study, Year, Total, sep=", "), method="REML")

# Generate Funnel Plots for supplemental
par(mfrow=c(1,2))
funnel(res, main="A")
funnel(res2, main="B")
```

**Figure 4: Funnel Plot of the Shannon Diversity and Bacteroidetes/Firmicutes Ratio Relative Risk.**  **A)** Overall there does not seem to be any bias associated with the studies selected for the Shannon Diversity analysis with studies falling on either side of the predicted value and those with smaller total n falling further away from this. **B)** For the Bacteroidetes/Firmicutes ratio there does not seem to be any associated bias. The overall pattern is similar to that observed for the Shannon Diversity analysis.
